---
title: "CleanedLeaflet"
format: html
---

Packages and data for first map 

```{r}
library(sf)
library(here)
library(tidyverse)
library(leaflet)
# I downloaded the data from this website and used the shp file 
# https://public-nps.opendata.arcgis.com/datasets/nps::nps-boundary-4/explore?location=37.015375%2C-81.906543%2C6.00
nps <- sf::read_sf("NPS_Land/nps_boundary.shp") |>
  sf::st_transform('+proj=longlat +datum=WGS84')

p_names <- nps|>
  mutate(popup = paste0('<a href =', nps$METADATA, '>',
                        nps$UNIT_NAME, '</a>'))
```

First Map 
```{r}
leaflet(nps) |>
  setView(lng = -110, lat = 39.833, zoom = 2.5) |>
  addTiles() |>
  addProviderTiles(providers$OpenStreetMap.Mapnik) |>
  addPolygons(color = "#006600",
              weight = 1.5,
              smoothFactor = .5,
              opacity = 1.0, 
              fillOpacity = 0.5,
              highlightOptions = highlightOptions(color = "#003300", weight = 2,
      bringToFront = TRUE),
              popup = p_names$popup)
```

Scrape for more info about National Park and clean it so I can merge with existing data 
```{r}
library(rvest)
library(stringr)

url <- "https://en.wikipedia.org/wiki/List_of_national_parks_of_the_United_States"

h <- read_html(url)

tab <- h |> html_nodes("table")

wiki_np_data <- tab[[1]] |> html_table()

wiki_np_clean <- wiki_np_data |> mutate(wiki_name = str_replace_all(Name, "[^[:alnum:]]", " ")) |>
  relocate(wiki_name)

nps_clean <- nps |> mutate(nps_name = str_remove_all(UNIT_NAME, " National") |>
                 str_remove_all(" Historical Site") |> 
                 str_remove_all( " Historical Park") |>
                 str_remove_all( " Historical Trail") |>
                 str_remove_all( " Park")) |>
  relocate(nps_name)
```

Merge with existing data for map 
```{r}
library(fuzzyjoin)
joined2 <- nps_clean |>
  stringdist_inner_join(wiki_np_clean, by = c("nps_name" = "wiki_name"), max_dist = 3) |>
  relocate(wiki_name, nps_name)

names_2 <- joined2 |>
  mutate(popup = paste0('<a href =', joined2$METADATA, '>',
                        joined2$UNIT_NAME, '</a>'))


leaflet(joined2) |>
  setView(lng = -98.583, lat = 39.833, zoom = 5) |>
  addTiles() |>
  addProviderTiles(providers$OpenStreetMap.Mapnik) |>
  addPolygons(color = "#006600",
              weight = 1.5,
              smoothFactor = .5,
              opacity = 1.0, 
              fillOpacity = 0.5,
              highlightOptions = highlightOptions(color = "#003300", weight = 2,
      bringToFront = TRUE),
      popup = names_2$popup)
```

Okay this has gotten me to the point where I can now plot all of the parks I have the extended info from wiki for. Now I need to figure out have to have that information pop up when the park is clicked on. I also want just the park name to appear when the mouse hovers over it. 
```{r}
joined_clean <- joined2 |> rename(Established = `Date established as park[12]`, Area = `Area (2023)[8]`, Visitors_2022 = `Recreation visitors (2022)[11]`) 

joined_clean <- joined_clean |> mutate(popup = paste(
  "Name: ", UNIT_NAME,
  "Date of Establishment: ", Established,
  "Visitors in 2022: ", Visitors_2022,
  "Description of Park: ", Description))

library(htmltools)
library(glue)

label_text <- glue(
  "<b> Name: </b> {joined_clean$UNIT_NAME}<br/>",
  "<b> Established: </b> {joined_clean$Established}<br/>",
  "<b> Visitors in 2022: </b> {joined_clean$Visitors_2022}<br/>",
  "<b> Description: </b> {joined_clean$Description}<br/>") |>
  lapply(htmltools::HTML)
  
leaflet(joined_clean) |>
  setView(lng = -110, lat = 39.833, zoom = 2.5) |>
  addTiles() |>
  addProviderTiles(providers$OpenStreetMap.Mapnik) |>
  addPolygons(color = "#006600",
              weight = 1.5,
              smoothFactor = .5,
              opacity = 1.0, 
              fillOpacity = 0.5,
              highlightOptions = highlightOptions(color = "#003300", weight = 2,
      bringToFront = TRUE),
      label = joined_clean$UNIT_NAME,
      popup = label_text)
```

Combined map 
```{r}
leaflet(nps) |>
  setView(lng = -110, lat = 39.833, zoom = 2.5) |>
  addTiles() |>
  addProviderTiles(providers$OpenStreetMap.Mapnik) |>
  addPolygons(color = "#09008a",
              weight = .5,
              smoothFactor = .5,
              opacity = 1.0, 
      label = nps$UNIT_NAME) |>
  addPolygons(color = "#006400",
              weight = 1.5,
              smoothFactor = .5,
              opacity = 1.0, 
              fillOpacity = 0.5,
              highlightOptions = highlightOptions(color = "#003b00", weight = 2,
      bringToFront = TRUE),
              label = joined_clean$UNIT_NAME,
              popup = label_text,
              data = joined_clean)
```

Okay to finish this app I am going to add data tables below the map, a select park(s) option on visitors plot, and another tab with a sheet of questions that can be answered using the app. (The Purpose of the project is going to be a basic educational app regarding national parks). 

Adding a data table to the bottom of the map 
```{r}
library(DT)

nps_display <- nps |> select(UNIT_NAME, UNIT_TYPE, STATE, METADATA )
datatable(nps_display)
joined_display <- joined_clean |> select(PARKNAME, STATE, Established, Area, Visitors_2022, Description)
joined_display <- joined_display |> rename(Name = PARKNAME, State = STATE, "Visitors in 2022" = Visitors_2022)

# to make this work for finding oldest/newest park I think it needs to be just the year it was established or turn data type into a year 
library(lubridate)

joined_display |> mutate(Date = mdy(Established)) |> view()

gsub("\\[.*?\\]", "", joined_display$Established)
joined_display <- joined_display |> mutate(Date = gsub("\\[.*?\\]", "", joined_display$Established)) |> mutate(Established_D = mdy(Date)) |> select(Name, State, Established_D, Area, `Visitors in 2022`, Description) |> rename(Established = Established_D)

datatable(joined_display) 
```






