---
title: "Question Sheet"
author: "Denali Stevens"
format: 
  html:
    output: false
    self-contained: true
---

# Section 1
Use the map and data tables to answer the following:

1. What is the oldest national park and when was it established?
```{r}
#| code-fold: true

# Yellowstone. It was established on March 1, 1872
```

2. What is the newest national park and when was it established?
```{r}
#| code-fold: true

# White Sands. It was established on December 20, 2019
```

3. What is the biggest national park by area? What state is it located in?
```{r}
#| code-fold: true

# Wrangell St. Elias located in AK
```

4. What is the smallest national park by area? What state is it located in?
```{r}
#| code-fold: true

# Gateway Arch located in IL
```

5. Which state has the most national parks? How many does it have 
```{r}
#| code-fold: true

# California has the most parks with a total of 9 in the state 
```

# Section 2
Use the collective visitors plot to answer the following:

1. Select 3 of your favorite parks
  a. Which one had the most visitors in 2000? How many did it have?
  b. How about the most visitors in 2016? Is it the same park as part a?

2. Can you figure out which park has the most visitors in this graphic? (Hint: you can sort the table by visitors to get an idea of which ones to look at)
```{r}
#| code-fold: true

# Great Smoky Mountains 
```

# Section 3

Use the single park visitors plot to answer the following:

1. Select Yellowstone National Park 
  a. What year did Yellowstone record the most visitors in this time period?
```{r}
#| code-fold: true

# In 2016, with 4,257,177 visitors 
```
  b. What year did it record the least visitors?
```{r}
#| code-fold: true

# In 2001, with 2,758,526
```
  
2. Now select your own park and answer the same questions!

  

